Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-08 paper/tv.tex
m

Now, the inclusions are disjunction-free formulas.

Next we provide the definitions and implementations of auxiliary functions \emph{Nullable}(\code{\delta}), \emph{First}(\code{fst}) and \emph{Derivative}(\code{D}) respectively.
Intuitively, the Nullable function \code{\delta(\effect)} returns a boolean value indicating whether \code{\effect} contains the empty trace; the First function \code{fst_{{\pi}}( \effect)} computes a set of possible initial time instances of \code{{\pi} : \effect}; and the Derivative function \code{D_{I\mysharp  t}^{{\pi}}(\effect)} computes a next-state effects after eliminating one event \code{I} w.r.t. the execution time \code{t} from the current effects \code{{\pi} : \effect}.

Next we provide the definitions and the key implementations\footnote{As 
the implementations according to basic regular expressions can be found 
in prior work \cite{keil2014symbolic}. 
%Here, we focus on presenting 
%the definitions and how do we deal with dependent values in the effects, 
%as the key novelties of this work. 
%More detail implementations can be found in the extended technical report\citeref{Demo}.
} of Nullable, First and Derivative respectively. Intuitively, 
the Nullable function \code{\delta_{{\pi}}(\es)} returns a boolean 
value indicating whether \code{{\pi} \wedge \es} contains the empty 
trace; the First function \code{fst_{{\pi}}( \es)} computes a set of 
possible initial events of \code{{\pi} \wedge \es}; and the Derivative 
function \code{D_{\anyevent{a}}^{{\pi}}(\es)} computes a next-state 
effects after eliminating one event \code{\anyevent{a}} from the current 
effects \code{{\pi} \wedge \es}.

@@

\begin{definition}[Nullable]\label{Nullable}
Given any timed sequence \code{\es} under condition \code{{\pi}}, we define \code{\delta_{{\pi}}(\es)} to be:\\
{
\code{\delta_{{\pi}}(\es) :bool {=}
\begin{cases}
      \code{true} & \code{if\ \empt \in \llbracket {\pi} \wedge \es \rrbracket_{\varphi}}\\
      \code{false} & \code{if\ \empt \notin \llbracket {\pi} \wedge \es \rrbracket_{\varphi}}
    \end{cases} }
    }, where\   
{ 
\begin{gather*}
\code{\delta_{\pi}(\bott) {=} false} 
\qquad\quad
\code{\delta_{\pi}(\empt) {=} true} 
\qquad\quad
\code{\delta_{\pi}(\ev){=} false}   
\qquad\quad
\code{\delta_{\pi}(\pi^\prime? \es){=} \delta_{\pi}(\es)}   
\qquad\quad\
    \code{\delta({\es^\star} ) {=} true}   
\\ 
\code{\delta_{\pi}(\es \seq \es_2) {=} \delta(\es_1) \wedge \delta(\es_2)}
\qquad\qquad\quad
  \code{\delta_{\pi}(\es_1 \choice \es_2) {=} \delta(\es_1) \vee \delta(\es_2)}   
  \\
\code{\delta_{\pi}(\es_1 || \es_2) {=} \delta(\es_1) \wedge \delta(\es_2)}
\qquad\qquad\quad
 \code{\delta_{\pi}(\es \mysharp  t) {=} SAT ({\pi} \wedge (t {=} 0) )}
\end{gather*}}
\end{definition}

@@

\begin{definition}[First]\label{First}
  Let \code{fst_{{\pi}}(\es): \{ (\ev,  t)  | 
  \pi {\wedge} t{\geq}0 \wedge (\ev \mysharp  t) {\cdot} \es_1 \in \llbracket  
  \pi \wedge \es \rrbracket \}} be the set of initial timed events  
  derivable from effects \code{\pi \wedge \es}. (\code{\llbracket  \pi \wedge
   \es \rrbracket } represents all the timed  traces contained in 
   \code{\pi \wedge \es}) 
  { 
   \begin{gather*} 
  \code{fst_{{\pi}}( \bott) {=} \{\} } \qquad\quad
  \code{fst_{{\pi}}(\empt) {=}  \{\} } \qquad\quad
  \code{fst_{{\pi}}( \ev) {=} \{ (\ev, t) \}  \ (t\ is\ fresh)} 
  \qquad\quad
     \code{fst_{{\pi}}(  \es \mysharp  t) {=}  fst_{\pi}(\es)}
   \\
  \code{fst_{{\pi}}( {\es^\star} ) {=} fst_{{\pi}}({ \es})}
  \qquad\ \ \  
  \qquad\quad
  \code{fst_{{\pi}}(  \es_1 {\vee} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }  \\
  \code{fst_{{\pi}}( \pi^\prime? \es ) {=} fst_{\pi}(\es)  }
      \qquad\qquad\quad
     \code{fst_{{\pi}}(  \es_1 {||} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }   
     \\
  {
  \code{fst_{{\pi}}(\es_1 {\seq} \es_2) {=} } 
  \begin{cases}
        \code{\code{fst_{{\pi}}(  \es_1)} \cup \code{fst_{{\pi}}(  \es_2)} } & \code{ if\ \delta(\es_1) {=} true}\\
        \code{\code{fst_{{\pi}}(  \es_1)}} & \code{  if\ \delta(\es_1) {=} false}
      \end{cases} 
      }  
  \end{gather*}
  }
\end{definition}

\begin{definition}[First-Tuples Unification]\label{unify} Given two first tuples \code{(I_1, \pi_1, t_1)} and \code{(I_2, \pi_2, t_2) }, we define that: 
%We deploy a \code{unify} function to merge two time instances, where \\ 
\code{unify\ (I_1, \pi_1, t_1)\  (I_2, \pi_2, t_2) = (I_1 \cup I_2, \pi_1 \wedge \pi_2 \wedge (t1 {=} t2), t_1)}.
\end{definition}

\begin{definition}[Instances Subsumption]\label{Subsumption}
Given two instances \code{I} and \code{J}, we define the subset relation \code{I {\subseteq} J} as: the set of present signals in \code{J} is a subset of the set of present signals in \code{I}, and the set of absent signals in \code{J} is a subset of the set of absent signals in \code{I}\footnote{As in having more constraints refers to a smaller set of satisfying instances.}.
Formally, 
 \begin{gather*} 
 \code{I {\subseteq} J \Leftrightarrow  \{\textbf{S}\ |\ (\textbf{S} {\mapsto} Present) \in J \}  \subseteq  \{\textbf{S}\ |\ (\textbf{S} {\mapsto} Present) \in I \}} \\
  \code{ \quad\  and  \ 
 \{\textbf{S}\ |\ (\textbf{S} \mapsto Absent) \in J \}  \subseteq  \{\textbf{S}\ |\ (\textbf{S} \mapsto Absent) \in I \}
 }
\end{gather*}

>>>>

Now, the inclusions are disjunction-free formulas. 
Next we provide the definitions and the key implementations\footnote{As 
the implementations according to basic regular expressions can be found 
in prior work \cite{keil2014symbolic}. 
%Here, we focus on presenting 
%the definitions and how do we deal with dependent values in the effects, 
%as the key novelties of this work. 
%More detail implementations can be found in the extended technical report\citeref{Demo}.
} of Nullable, First and Derivative respectively. Intuitively, 
the Nullable function \code{\delta_{{\pi}}(\es)} returns a boolean 
value indicating whether \code{{\pi} \wedge \es} contains the empty 
trace; the First function \code{fst_{{\pi}}( \es)} computes a set of 
 initial \emph{heads}, denoted as \code{h},  of \code{{\pi} \wedge \es}; and the Derivative 
function \code{D_{h}^{{\pi}}(\es)} computes a next-state 
effects after eliminating the head  \code{h} from the current 
effects \code{{\pi} \wedge \es}.

@@

\begin{definition}[Nullable]\label{Nullable}
Given any \code{\effect {=} \pi {\wedge} \es}, %we define 
%\code{\delta_{{\pi}}(\es)} to be \\
{
  \code{\delta_{{\pi}}(\es) :bool {=}
  \begin{cases}
        \code{true} & \code{if\ \empt \in \llbracket {\pi} {\wedge} \es \rrbracket_{\varphi}}\\
        \code{false} & \code{if\ \empt \notin \llbracket {\pi} {\wedge} \es \rrbracket_{\varphi}}
      \end{cases} }
      }, where:  
\begin{comment}
 a boolean value, indicating whether \code{\effect} contains the empty trace, i.e., $\epsilon$. 
We define recursively:  
\end{comment} 
{ 
\begin{gather*}
\code{\delta_{\pi}(\bott) {=} false} 
\qquad\quad
\code{\delta_{\pi}(\empt) {=} true} 
\qquad\quad
\code{\delta_{\pi}(\ev){=} false}   
\qquad\quad
\code{\delta_{\pi}(\pi^\prime? \es){=} \delta_{\pi}(\es)}   
\qquad\quad\
    \code{\delta({\es^\star} ) {=} true}   
\\ 
\code{\delta_{\pi}(\es \seq \es_2) {=} \delta(\es_1) \wedge \delta(\es_2)}
\qquad\qquad\quad
  \code{\delta_{\pi}(\es_1 \choice \es_2) {=} \delta(\es_1) \vee \delta(\es_2)}   
  \\
\code{\delta_{\pi}(\es_1 || \es_2) {=} \delta(\es_1) \wedge \delta(\es_2)}
\qquad\qquad\quad
 \code{\delta_{\pi}(\es \mysharp  t) {=} SAT ({\pi} \wedge (t {=} 0) )}
\end{gather*}}
\end{definition}

@@

\begin{definition}[Heads]\label{Heads}
If \code{h} is a head of \code{\pi \wedge \es}, then there exists 
\code{\pi^\prime} and \code{\es^\prime}, 
such that \code{\pi \wedge \es} {=} \code{\pi^\prime \wedge h \cdot \es^\prime}. 
   %We adopt the following naming convention for transition labels: 
A head can be  $t$, denoting a pure time passing; 
  $\ev $, denoting a pure event passing; 
  $(\ev, t)$, denoting an event passing which takes time \code{t}. 
\end{definition}

\begin{definition}[First]\label{First}
  Given any \code{\effect {=} \pi \wedge \es},
  \code{fst_{{\pi}}(\es)} return a set of heads, be the set of initial elements   
  derivable from effects \code{\pi \wedge \es}, where: 
    { 
   \begin{gather*} 
  \code{fst_{{\pi}}( \bott) {=} \{\} } \qquad\ \ 
  \code{fst_{{\pi}}(\empt) {=}  \{\} } \qquad\ \ 
  \code{fst_{{\pi}}( \ev) {=} \{ \ev \}  }
  \qquad\ \ 
  \code{fst_{{\pi}}(  \epsilon \mysharp  t) {=}  \{ t \}}
  \qquad\ \ 
  \code{fst_{{\pi}}( {\es^\star} ) {=} fst_{{\pi}}({ \es})}
   \\
   \code{fst_{{\pi}}(\es \mysharp  t) {=}  \{(\ev, t) \ | \ \ev \in fst_{\pi}(\es) \}} 
   \ \ \  \emph{(t\ is\ fresh)}
  \qquad\quad 
  \qquad\quad
  \code{fst_{{\pi}}(  \es_1 {\vee} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }  \\
  \code{fst_{{\pi}}( \pi^\prime? \es ) {=} fst_{\pi}(\es)  }
      \qquad\qquad\quad
     \code{fst_{{\pi}}(  \es_1 {||} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }   
     \\
  {
  \code{fst_{{\pi}}(\es_1 {\seq} \es_2) {=} } 
  \begin{cases}
        \code{\code{fst_{{\pi}}(  \es_1)} \cup \code{fst_{{\pi}}(  \es_2)} } & \code{ if\ \delta(\es_1) {=} true}\\
        \code{\code{fst_{{\pi}}(  \es_1)}} & \code{  if\ \delta(\es_1) {=} false}
      \end{cases} 
      }  
  \end{gather*}
  }
\end{definition}

\begin{definition}[Timed Event Subsumption]\label{Subsumption}
Given two instances \code{I} and \code{J}, we define the subset relation \code{I {\subseteq} J} as: the set of present signals in \code{J} is a subset of the set of present signals in \code{I}, and the set of absent signals in \code{J} is a subset of the set of absent signals in \code{I}\footnote{As in having more constraints refers to a smaller set of satisfying instances.}.
Formally, 
 \begin{gather*} 
 \code{I {\subseteq} J \Leftrightarrow  \{\textbf{S}\ |\ (\textbf{S} {\mapsto} Present) \in J \}  \subseteq  \{\textbf{S}\ |\ (\textbf{S} {\mapsto} Present) \in I \}} \\
  \code{ \quad\  and  \ 
 \{\textbf{S}\ |\ (\textbf{S} \mapsto Absent) \in J \}  \subseteq  \{\textbf{S}\ |\ (\textbf{S} \mapsto Absent) \in I \}
 }
\end{gather*}
